Assessment and implementation of resource road simulation methods for projection of cumulative effects of disturbance on wildlife

### 2023-04-11 analysis/paper/paper.Rmd
add Hardy QGIS to proj methods fig

The DLCP and MST methods connect target or landing locations within each cutblock to the existing road network, so an important first step is to select one or more landing locations in each block. We investigated the sensitivity of results to the pattern and density of landing locations by comparing results derived from the centroid of each cutblock to regular or randomly selected landing densities of 1 (low) and 10 (high) points per km^2^ (Figure \@ref(fig:proj-meth-fig)). We used default settings in the Hardy QGIS plugin except for kidding distance which was set to 200m. A skidding distance of 200 m corresponds to \~ 9 pts/km^2^. We compared two resolutions of the cost surface by aggregating the 1 ha resolution cost raster to a coarser 100 ha (1000 by 1000 m) resolution. In total, we compared 12 road projection method variants to a cutblocks only scenario with no projected roads (Table \@ref(tab:method-tbl)).

@@

sampleDens <- c(low,high,low,high,low)
sampleType <- c("regular","regular","random","random","centroid")
paramTable <- tibble(sampleType, sampleDens,
                     method = "mst",
                     runTime = vector("list", length(sampleDens)),
                     output = vector("list", length(sampleDens)),
                     roadDisturbance = vector("list", length(sampleDens)),
                     roadDensity = vector("list", length(sampleDens)),
                     roadPresence = vector("list", length(sampleDens)),
                     distanceToRoad = vector("list", length(sampleDens)),
                     forestryDisturbance = vector("list", length(sampleDens))) %>%
  distinct()

@@

lcpProj <- projectAll(tsbs = tsb,
                       paramTable = paramTable %>% 
                        filter(sampleType == "centroid") %>% 
                        mutate(method = "lcp"),
                       costSurface = tsaCost,
                       cutblocks = cutblocks,
                       existingRoads = exRoads,
                       fileLocation = here(data_path_drvd, "..", "for_fig"))

allProj <- bind_rows(mstProj, dlcpProj, lcpProj, .id = "method") %>%
  arrange(desc(sampleType), sampleDens)

allMaps <- purrr::map(
  1:nrow(allProj),
  ~ qtm(tsaCost, raster.style = "cont", raster.palette = "Greys",
        raster.alpha = 0.5)+
    qtm(cutblocks, fill = NULL, borders = "grey40", borders.lwd = 1.5)+
    qtm(roads, lines.col = "red", lines.lwd = 2)+
    qtm(read_sf(here(allProj$output[[.x]])), lines.col = "yellow", lines.lwd = 2)+
    qtm(exRoads, lines.col = "black", lines.lwd = 2)+
    tm_layout(legend.show = FALSE,
              main.title = paste(ifelse(allProj$method[.x]== 1, "MST",
                                        ifelse(allProj$method[.x]== 2, "DLCP", "LCP")),
                                 allProj$sampleType[.x],
                                 ifelse(allProj$sampleType[.x] == "centroid", "",
                                        ifelse(allProj$sampleDens[.x] == 1e-06, "low density", "high density"))),
              main.title.size = 0.75))

tmap_save(tmap_arrange(allMaps, ncol = 2),
          here("analysis/figures/projection_methods_figure.png"),
          dpi = 300, height = 8, width = 3.543)
```

```{r proj-meth-fig, fig.cap="We compared minimum spanning tree (MST), dynamic least cost paths (DLCP) and least cost path (LCP) road network projection algorithms. Landing locations were selected as the centroid of the cutblock, or a random or regularly spaced sample of points inside each cutblock (Table 2). Target cutblocks are outlined in grey, existing roads are black, projected roads are yellow, observed roads are red. Cost (1 ha resolution) is indicated by grey shading."}
knitr::include_graphics(here("analysis/figures/projection_methods_figure.png"))
```

>>>>

The DLCP and MST methods connect target or landing locations within each cutblock to the existing road network, so an important first step is to select one or more landing locations in each block. We investigated the sensitivity of results to the pattern and density of landing locations by comparing results derived from the centroid of each cutblock to regular or randomly selected landing densities of 1 (low) and 10 (high) points per km^2^ (Figure \@ref(fig:proj-meth-fig)). We used default settings in the Hardy QGIS plugin except for skidding distance which was set to 200m. A skidding distance of 200 m corresponds to \~ 9 pts/km^2^. We compared two resolutions of the cost surface by aggregating the 1 ha resolution cost raster to a coarser 100 ha (1000 by 1000 m) resolution. In total, we compared 12 road projection method variants to a cutblocks only scenario with no projected roads (Table \@ref(tab:method-tbl)).

@@

sampleDens <- c(high, high, low)
sampleType <- c("regular","random","centroid")
paramTable <- tibble(sampleType, sampleDens,
                     method = "mst",
                     runTime = vector("list", length(sampleDens)),
                     output = vector("list", length(sampleDens)),
                     roadDisturbance = vector("list", length(sampleDens)),
                     roadDensity = vector("list", length(sampleDens)),
                     roadPresence = vector("list", length(sampleDens)),
                     distanceToRoad = vector("list", length(sampleDens)),
                     forestryDisturbance = vector("list", length(sampleDens))) %>%
  distinct()

@@

allProj <- bind_rows(mstProj, dlcpProj, .id = "method") %>%
  arrange(desc(sampleType), sampleDens)

# add Hardy QGIS results for the same area

allProj <- bind_rows(
  allProj,
  tibble(sampleType = "", sampleDens = 0, method = "Hardy QGIS",
             output = list(here(data_path_drvd, "..", "for_fig", 
                                "HardyQGISTest.gpkg")))) %>% 
  mutate(mapTitle =  paste(ifelse(method == 1, "MST",
                                  ifelse(method == 2, "DLCP", method)),
                           sampleType,
                           ifelse(sampleDens == 1e-06 & 
                                    sampleType != "centroid", "low density", 
                                  ifelse(sampleDens == 1e-05, 
                                         "high density", ""))))

# add extra row for legend
allProj <- bind_rows(allProj, 
                     slice(allProj, 1))

allMaps <- purrr::map(
  1:nrow(allProj),
  ~ qtm(tsaCost, raster.style = "cont", raster.palette = "Greys",
        raster.alpha = 0.5, raster.title = "Cost")+
    qtm(cutblocks, fill = NULL, borders = "grey40", borders.lwd = 1.5)+
    qtm(roads, lines.col = "red", lines.lwd = 2)+
    qtm(read_sf(here(allProj$output[[.x]])), lines.col = "yellow", lines.lwd = 2)+
    qtm(exRoads, lines.col = "black", lines.lwd = 2)+
    tm_layout(legend.show = .x == nrow(allProj),
              legend.only = .x == nrow(allProj),
              main.title = allProj$mapTitle[[.x]],
              main.title.size = 0.75))

allMaps[[nrow(allProj)]] <- allMaps[[nrow(allProj)]]+
  tm_add_legend(type = "line", 
                labels = c("Existing roads", "Projected roads", "Observed roads"),
                col = c("black", "yellow", "red"))+
    tm_add_legend(type = "fill", 
                labels = "Cutblocks",
                col = NA, border.col = "grey40")+
  tm_layout(legend.only = TRUE, legend.position = c("center", "center"))

tmap_save(tmap_arrange(allMaps, ncol = 3),
          here("analysis/figures/projection_methods_figure.png"),
          dpi = 300, height = 5, width = fig_widths["two"])
```

```{r proj-meth-fig, fig.cap="We compared minimum spanning tree (MST), dynamic least cost paths (DLCP) and the Hardy QGIS plugin road network projection algorithms. Landing locations were selected as the centroid of the cutblock, or a random or regularly spaced sample of points inside each cutblock (Table 2). Target cutblocks are outlined in grey, existing roads are black, projected roads are yellow, observed roads are red. Cost (1 ha resolution) is indicated by grey shading, with darker grey indicating higher cost. The large four sided cutblock in the top right was added to give a better example of the algorithms behaviour inside cutblocks."}
knitr::include_graphics(here("analysis/figures/projection_methods_figure.png"))
```
